Post-Disturbance Tree Community Trajectories in a Neotropical Forest

### 2018-04-05 WholePlotTrajectories.tex
Scripts

\JournalInfo{Hal 00679993} % Journal information
\Archive{DOI xxxx} % Additional notes (e.g. copyright, DOI, review/research article)

@@

@@

\subsection{Inventories protocol and dataset
correction}\label{inventories-protocol-and-dataset-correction}

The study site corresponds to a tropical rainforest with a dominance of
Fabaceae, Chrysobalanaceae, Lecythidaceae and Sapotaceae botanical
families. In the twelve experimental plots of the experiment, all trees
above 10 cm DBH were mapped and measured annually since 1984. During
inventories, trees were first identified with a vernacular name assigned
by the field team, and afterward with a scientific name assigned by a
botanist during regular botanical campaigns. In 1984, specific
vernacular names were given to 62 commercial or common species whereas
other less common species were identified under two identifiers only
separating trees and palm trees. The botanical campaigns carried every 5
to 6 years to identify all trees at the species level only started in
2003 and identification practices varied among plots and successive
campaigns. This raised methodological issues as vernacular names usually
correspond to different botanical species, resulting in significant
taxonomic uncertainties that had to be propagated to composition and
diversity metrics throught a Bayesian framework. The uncertainty
propagation was done by the replenishment of inventories completed at
genus level from real incomplete ones on the basisof
vernacular/botanical names association. Vernacular names were replaced
through multinomial trials
(M\textsubscript{v}({[}\emph{s\textsubscript{1}, s\textsubscript{2},
\ldots{}, s\textsubscript{N}}{]} ,{[}\emph{\(\alpha\)\textsubscript{1},
\(\alpha\)\textsubscript{2},\ldots{}, \(\alpha\)\textsubscript{N}}{]}) )
based on the observed association probability
{[}\emph{\(\alpha\)\textsubscript{1},
\(\alpha\)\textsubscript{2},\ldots{}, \(\alpha\)\textsubscript{N}}{]}
between each vernacular name \emph{v} and the species
{[}\emph{s\textsubscript{1}, s\textsubscript{2}, \ldots{},
s\textsubscript{N}}{]} recorded in the inventory. To avoid remaining
identification caveats and consider complete inventories, the simulated
inventories were then reported at genus. See appendix 1 and
\citet{Aubry-Kientz2013} for the detailed methodology.

The functional diversity metrics used a dataset for 6 functional traits
representing leaf economics (leaves thickness, toughness, total
chlorophyll content and specific leaf area, the leaf area per unit dry
mass) and wood economics (wood specific gravity and bark thickness), and
life history traits (maximum specific height and seed mass). The trait
database came from the BRIDGE project \footnote{http://www.ecofog.gf/Bridge/index.html}
where traits values were assessed from a selection of individuals
located in nine permanent plots in French Guiana, including two in
Paracou. Missing trait values were filled using multivariate imputation
by chained equation (mice) restricted to samples pertaining to the next
higher taxonomic level, in order to account for the phylogenetic signal
of the functional traits. The dataset comprised 294 botanical species
pertaining to 157 botanical genera. Whenever a species inventoried was
not in the dataset, it was attributed a set of traits values randomly
sampled among species of the same next higher taxonomic level. As seed
mass information corresponds to a classification into mass classes, no
data filling process was applied so analysis were performed considering
the 414 botanical species of the seed mass dataset. All composition and
diversity metrics corresponded to the average obtained after 50
iterations of the taxonomic uncertainty propagation framework and of the
filling process of missing trait values.

@@

The taxonomic diversity was assessed through Richness and the Hill
number translation of Shannon and Simpson indices \citep{Hill1973}. To
tackle the unequal number of recruited trees among treatments the
indices bias corrected estimator were used, following
\citep{chao2015estimating, Marcon2015b}. These three indices belong to
the set of HCDT or generalized entropy, respectively corresponding to
the 0, 1 and 2 order of diversity (q), which proved well suited for
diversity studies \citep{Patil1982, Tothmeresz1995}. The functional
diversity was reported using the Rao index of quadratic entropy which
combines species abundance distribution and average pairwise
dissimilarity based on all functional and life traits.

@@

\emph{The distance from the initial condition was significantly
time-dependent (P\textless{}0.01).} \emph{The coordinates of the
functional traits were measured in the two-dimensional ordination space
that mapped plots evolution along time (graphs not showed) which
revealed that the trajectories of disturbed plots along time headed
toward acquisitive functional strategies (from high SWG to high SLA and
chlorophyll content).}

@@

In line with the assumtions of the IDH the observed diversity
trajectories confirmed the increase of communities evenness and
functional richness after disturbance. The consistency of the IDH has
been debated for the higly diverse tropical forests, some observing the
importance of environmental filters to shape communities assembly and
others arguing for a genuine stochastic recovery
\citep{Hubbell1999, Molino2001}. Consistently to the trajectories drawn
here, a previous study of the same experiment revealed the directional
changes in communities composition reflecting environmental filtering
after disturbance \citep{Molino2001}. Disturbance regimes brought a
unimodal response depending on disturbance intensity as suggested by the
outcome of 10 years and 20 years post-disturbance
analysis\citep{Molino2001, Baraloto2012a}. As observed after logging in
a Boranean rainforest \citep{Cannon1998} or after cyclones in the
Solomon islands \citep{Burslem2000}, communities diversity showed a
unimodal response to disturbance and a maintenance of diversity levels,
except for one of the intensively disturbed community which evenness was
still increasing (Figure \ref{fig:Fig5}). Despite the recovery of
communities diversity, though, taxonomic composition was clearly altered
(Figure \ref{fig:Fig3}) implying a long-term, if not permenant,
composition change after disturbance as observed in East African
tropical forests \citep{Bonnell2011}.

@@

The diversity and composition functional trajectories showed a
decoupling between functional and taxonomic after the most intense
disturbance (treatments T2 and T3). Plots functional characteristic
returned to their initial values, similar to those of control plots,
faster than their taxonomic counterparts (Figures \ref{fig:Fig2} to
\ref{fig:Fig4}). As functional traits are the most direct link between
biodiversity and ecosystem functioning \citep{Diaz2005} the recovery of
communities functional characteristics would go along the recovery of
ecosystems processes although taxonomic composition remain altered
\citep{Guariguata2001}. Such decoupling suggested the consistency of
functional overlap and maybe functional versatility, implying a variety
of functioning from the same characteristic, in the pre-disturbance
condition. Functional overlap has been observed in highly diverse
ecosystems \citep{Bellwood2006} and suggested as one mechanisms for high
diversity maintenance. Knowing the decoupling between taxonomic and
functional trajectories, this functional overlap may not be recovering
after disturbance and therefore radically modify communities resilience
and resistance \citep{Trenbath1999, Elmqvist2003, Diaz2005}. Given the
impaired functional overlap, the persitant compositional changes
regarding disturbance resistant species \citep{Haddad2008} and probably
lianas or epiphytes \citep{Martin2013}, and the modification of
environmental conditions like soils nutrient cycling and compaction
\citep{Olander2005} the entirety of communities recovery after intense
disturbance is deeply questioned \citep{Chazdon2003a}. New conditions
would not only be longer lasting but self-maintained as tied to
disturbance regime \citep{Burslem2000}. Although communities recovered
the core of their initial functioning, composition seemed permanently
altered after intense disturbance. Specificaly, this would impair
species contingent to undisturbed forests, threatening their
maintenance, and run the risk to loose cornerstone species and trigger
unexpected ecological consequences \citep{Diaz2005, Gardner2007}.

@@

\bibliographystyle{mee}
\makeatletter
% The filename has .bib extension the must be eliminated
\filename@parse{references.bib}
% parse stores the file name in base. Extension starts at the first dot, so don't use dots in file names.
\bibliography{\filename@base}
\makeatother

>>>>

\JournalInfo{Hal xxx} % Journal information
\Archive{DOI xxxx} % Additional notes (e.g. copyright, DOI, review/research article)

@@

\begin{verbatim}
## Warning: package 'kableExtra' was built under R version 3.3.3
\end{verbatim}

\begin{verbatim}
## Warning: package 'shape' was built under R version 3.3.3
\end{verbatim}

@@

\subsection{Inventories protocol and dataset
collection}\label{inventories-protocol-and-dataset-collection}

The study site corresponds to a tropical rainforest with a dominance of
Fabaceae, Chrysobalanaceae, Lecythidaceae and Sapotaceae botanical
families. In the twelve experimental plots of the experiment, all trees
above 10 cm DBH were mapped and measured annually since 1984. During
inventories, trees were first identified with a vernacular name assigned
by the field team, and afterward with a scientific name assigned by a
botanist during regular botanical campaigns. In 1984, specific
vernacular names were given to 62 commercial or common species whereas
other less common species were identified under two identifiers only
separating trees and palm trees. The botanical campaigns carried every 5
to 6 years to identify all trees at the species level only started in
2003 and identification practices varied among plots and successive
campaigns. This raised methodological issues as vernacular names usually
correspond to different botanical species, resulting in significant
taxonomic uncertainties that had to be propagated to composition and
diversity metrics throught a Bayesian framework. The uncertainty
propagation was done by the replenishment of inventories completed at
genus level from real incomplete ones on the basisof
vernacular/botanical names association.

Vernacular names were replaced through multinomial trials
\(M_v\Big(\big[s_1, s_2, …, s_N\big],\big[\alpha_1, \alpha_2,…, \alpha_3\big]\Big)\)
based on the observed association probability
\(\big[\alpha_1, \alpha_2,…, \alpha_3\big]\) between each vernacular
name \emph{v} and the species \(\big[s_1, s_2, …, s_N\big]\) recorded in
the inventory.

To avoid remaining identification caveats and consider complete
inventories, the simulated inventories were then reported at genus. See
appendix 1 and \citet{Aubry-Kientz2013} for the detailed methodology.

The functional diversity metrics used a dataset for 6 functional traits
representing leaf economics (leaves thickness, toughness, total
chlorophyll content and specific leaf area, the leaf area per unit dry
mass) and wood economics (wood specific gravity and bark thickness), and
life history traits (maximum specific height and seed mass).

The trait database came from the BRIDGE project {[}\^{}1{]} where traits
values were assessed from a selection of individuals located in nine
permanent plots in French Guiana, including two in Paracou. Missing
trait values were filled using multivariate imputation by chained
equation (mice) restricted to samples pertaining to the next higher
taxonomic level, in order to account for the phylogenetic signal of the
functional traits. The dataset comprised 294 botanical species
pertaining to 157 botanical genera.

Whenever a species inventoried was not in the dataset, it was attributed
a set of traits values randomly sampled among species of the same next
higher taxonomic level. As seed mass information corresponds to a
classification into mass classes, no data filling process was applied so
analysis were performed considering the 414 botanical species of the
seed mass dataset. All composition and diversity metrics corresponded to
the average obtained after 50 iterations of the taxonomic uncertainty
propagation framework and of the filling process of missing trait
values.

{[}\^{}1{]} http://www.ecofog.gf/Bridge/

@@

The taxonomic diversity was assessed through Richness and the Hill
number translation of Shannon and Simpson indices \citep{Hill1973}. To
tackle the unequal number of recruited trees among treatments the
indices bias corrected estimator were used, following
\citep{Chao2015, Marcon2015b}. These three indices belong to the set of
HCDT or generalized entropy, respectively corresponding to the 0, 1 and
2 order of diversity (q), which proved well suited for diversity studies
\citep{Patil1982, Tothmeresz1995}. The functional diversity was reported
using the Rao index of quadratic entropy which combines species
abundance distribution and average pairwise dissimilarity based on all
functional and life traits.

@@

\begin{quote}
The distance from the initial condition was significantly time-dependent
(P\textless{}0.01). The coordinates of the functional traits were
measured in the two-dimensional ordination space that mapped plots
evolution along time (graphs not showed) which revealed that the
trajectories of disturbed plots along time headed toward acquisitive
functional strategies (from high SWG to high SLA and chlorophyll
content).
\end{quote}

@@

In line with the assumtions of the IDH the observed diversity
trajectories confirmed the increase of communities evenness and
functional richness after disturbance.
\textless{}\textless{}\textless{}\textless{}\textless{}\textless{}\textless{}
HEAD The consistency of the IDH has been debated for the higly diverse
tropical forests, some observing the importance of environmental filters
to shape communities assembly and others arguing for a genuine
stochastic recovery. Consistently to the trajectories drawn here, a
previous study of the same experiment revealed the directional changes
in communities composition reflecting environmental filtering after
disturbance. Disturbance regimes brought a unimodal response depending
on disturbance intensity as suggested by the outcome of 10 years and 20
years post-disturbance
analysis\citep{Hubbell1999, Molino2001, Baraloto2012a}. As observed
after logging in a Boranean rainforest \citep{Cannon1998} or after
cyclones in the Solomon islands \citep{Burslem2000}, communities
diversity showed a unimodal response to disturbance and a maintenance of
diversity levels, except for one of the intensively disturbed community
which evenness was still increasing \ref{fig:Fig5}. Despite the recovery
of communities diversity, though, taxonomic composition was clearly
altered \ref{fig:Fig3} implying a long-term, if not permenant,
composition change after disturbance as observed in East African
tropical forests \citep{Bonnell2011}.

@@

The diversity and composition functional trajectories showed a
decoupling between functional and taxonomic after the most intense
disturbance (treatments T2 and T3). Plots functional characteristic
returned to their initial values, similar to those of control plots,
faster than their taxonomic counterparts (Figures \ref{fig:Fig2} to
\ref{fig:Fig4}). As functional traits are the most direct link between
biodiversity and ecosystem functioning \citep{Diaz2005} the recovery of
communities functional characteristics would go along the recovery of
ecosystems processes although taxonomic composition remain altered
\citep{Guariguata2001}. Such decoupling suggested the consistency of
functional overlap and maybe functional versatility, implying a variety
of functioning from the same characteristic, in the pre-disturbance
condition. Functional overlap has been observed in highly diverse
ecosystems \citep{Bellwood2006} and suggested as one mechanisms for high
diversity maintenance. Knowing the decoupling between taxonomic and
functional trajectories, this functional overlap may not be recovering
after disturbance and therefore radically modify communities resilience
and resistance \citep{Trenbath1999, Elmqvist2003, Diaz2005}. Given the
impaired functional overlap, the persitant compositional changes
regarding disturbance resistant species \citep{Haddad2008} and probably
lianas or epiphytes \citep{Martin2013}, and the modification of
environmental conditions like soils nutrient cycling and compaction
\citep{Olander2005} the entirety of communities recovery after intense
disturbance is deeply questioned \citep{Chazdon2003a}. New conditions
would not only be longer lasting but self-maintained as tied to
disturbance regime \citep{Burslem2000}. Although communities recovered
the core of their initial functioning, composition seemed permanently
altered after intense disturbance. Specificaly, this would impair
species contingent to undisturbed forests, threatening their
maintenance, and run the risk to loose cornerstone species and trigger
unexpected ecological consequences
\citep{Jones1994, Diaz2005, Gardner2007}.

@@

\bibliographystyle{mee}
\bibliography{references.bib}
